Spontaneous Similarity Judgments of Musical Tones played with various Playing Techniques

### 2017-12-19 paper/lagrangeSpontaneousSimilarity2017.tex
adding data

In order to model the early stages of the mamalian auditory system, the time scattering transform \cite{anden2014deep}, obtained by applying a one dimensional
wavelet transform in time to a time-frequency wavelet scalogram. This transform thus encodes well modulations over time of energy distributed over frequency, dimensions that are shown to be perceptually important \cite{dau1997modeling}.

@@

Those methods are designed for considering only one reference grouping. Though, in our case, there is one reference grouping per subject. One approach is to find one consensus grouping that maximally represent the 31 reference groupings. In this paper, the chosen consensus grouping is taken from 3 solutions issued from 3 different algorithmic approach. This consensus is chosen as the one that has the best averaged NMI when compared to the 31 reference groupings \cite{strehl2002cluster}.

@@

The performance of the different methods are evaluated in a ranking setting. The well known precision at rank $k$ is thus considered, with $k=5$.

@@

The time scattering is computed over a range frame sizes, from 25 ms to 1 second. The Q factor is set to 8 in order to obtain a frequency decomposition that is equivalent to the mel scale. As the frame increase, the dimensionality increases. The resulting scattering features are then logarithmically compressed and normalized using the median computed over the dataset.

@@

For the sake of reproducibility, the code of the proposed method and the experimental protocol is available\footnote{https://github.com/mathieulagrange/paperSpontaneousSimilarity}. For convenience, the features (scattering and mfccs) are alos provided, but the software can be used to process from scratch the SOL database.

>>>>

In order to model the early stages of the mamalian auditory system, the time scattering transform (TST) \cite{anden2014deep}, obtained by applying a one dimensional
wavelet transform in time to a time-frequency wavelet scalogram. This transform thus encodes well modulations over time of energy distributed over frequency, dimensions that are shown to be perceptually important \cite{dau1997modeling}.

@@

Those methods are designed for considering only one reference grouping. Though, in our case, there is one reference grouping per subject. One approach is to find one consensus grouping that maximally represent the 31 reference groupings. In this paper, the chosen consensus grouping is taken from 3 solutions issued from 3 different algorithmic approaches. This consensus is chosen as the one that has the best averaged NMI when compared to the 31 reference groupings \cite{strehl2002cluster}.

@@

The performance of the different methods are evaluated in a ranking setting. The precision at rank $k$ is thus considered, with $k=5$. For one sound recording taken as the query, the $k$ nearest sound recordings are retrieved from the remaining items of the database. The precision is the number of neighbors that have the same label as the query in the grouping we consider as reference divided by $k$. This process is repeated for each sound recordings taken as query, and the performance is averaged over each query. For example, if a sound is played by a Violin, and among the 5 nearest neighbors of this query sound, 2 are played by a Violin, the precision is $0.4$.

When considering several groupings as reference, the precision is averaged per groupings. In our case, the precision is thus computed for the 31 reference groupings and averaged.

@@

The time scattering is computed over a range frame sizes, from 25 ms to 1 second. The Q factor is set to 8 in order to obtain a frequency decomposition that is equivalent to the mel scale. As the frame size increase, the dimensionality increases. Unless otherwise stated, the frame size is one second. The resulting scattering features are then logarithmically compressed and normalized using the median computed over the dataset.

@@

In order to validate the proposed model, some experiments are performed using the above described methods. 3 experimental factors with 2 or 3 modalities are considered:
\begin{itemize}
  \item type of features: MFCCs or TST
  \item type of projection: LDA, LMNN or no projection
  \item type of reference for the optimization of the projection if any: consensus (single consensus grouping) or reference (the 31 reference groupings)
\end{itemize}

\begin{table}
\begin{tabular}{lcc}
  MFCCs & - & - &   85.07 \pm 6.19 \\
  MFCCs & LDA &  consensus  &   81.50 \pm 7.65 \\
MFCCs & LMNN & consensus &   86.22 \pm 5.92 \\
MFCCs & LMNN & reference & 86.31 \pm 5.91 \\
% MFCCs & LMNN & reference &   86.18 \pm 6.05 \\
TST & - & - &   87.01 \pm 5.81 \\
TST & LDA & consensus &   80.95 \pm 10.37 \\
TST & LMNN  & reference  &   93.31 \pm 3.92 \\
TST & LMNN & consensus &   94.80 \pm 3.26 \\
% TST & LMNN & reference &   98.09 \pm 1.28 \\
\end{tabular}
\end{table}

separate judgments another table

For the sake of reproducibility, the code of the proposed method and the experimental protocol is available\footnote{https://github.com/mathieulagrange/paperSpontaneousSimilarity}. For convenience, the features (scattering and mfccs) are also provided, but the software can be used to process audio data.
